---
title: "Exploratory plots"
author: "Ruby Zhang"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
rm(list = ls())
```

# Load libraries
```{r}
library(tidyverse)
library(GGally) # ggpairs()
library(corrplot) # corrplot()
library(gridExtra) # grid.arrange()
library(tidymodels) # initial_split()
library(ggplot2) # gm_scatterplot
library(glmnet) # glmnet()
```

# Load data
Notes: 
- track_id unique
- track_name not unique 
(keep different versions by different singers, rid of different versions by the same singer)
```{r}
data_jenny <- read.csv("unique_tracks_genres.csv")
final_data <- select(data_jenny, -X, -track_id, -artists, -album_name, -track_name, -track_genre)

# convert into factors
final_data$explicit <- as.numeric(as.factor(final_data$explicit))-1 # 0 for FALSE, 1 for TRUE
final_data$mode <- as.integer(final_data$mode) # 0 for minor, 1 for major

# create new dummy showing if track has >1 genre
genres <- c("pop", "rock", "country", "jazz", "electronic", "classical", "world", "kids", "other", "rap")
final_data$two_genre <- as.numeric(rowSums(final_data[, genres]) == 2)
final_data$three_genre <- as.numeric(rowSums(final_data[, genres]) == 3)
final_data$four_genre <- as.numeric(rowSums(final_data[, genres]) == 4)
final_data$five_genre <- as.numeric(rowSums(final_data[, genres]) == 5)
```

# Exporatary analysis

We can get a summary of the variables of interest. See visual representation later.
```{r}
summary(final_data)
```


## Distribution of variables wrt popularity
```{r}
# Histogram
col_names <-names(final_data)
for (i in seq_along(col_names)){
  hist(final_data[,i], main=paste("Histogram of", col_names[[i]]))
}
```

```{r}
# Load the ggplot2 library
library(ggplot2)

# Generate some sample data
data <- data.frame(x_values = rnorm(1000, mean = 0.5, sd = 0.1),
                   y_values = rnorm(1000, mean = 0.95, sd = 0.1),
                   variable_of_interest = rnorm(1000, mean = 2, sd = 0.1))

# Create a histogram with color gradient and color coding for variable of interest
ggplot(data, aes(x = x_values, fill = ..density.., color = variable_of_interest)) + 
  geom_histogram(binwidth = 0.01, color = "white", aes(y = ..density..)) +
  scale_fill_gradient(low = "blue", high = "red") +
  scale_color_gradient(low = "green", high = "red") +
  theme_classic() +
  labs(title = "Histogram with Color Gradient and Color Coding",
       x = "X Values", y = "Density", color = "Variable of Interest")
```

## Correlation plot
```{r}
final_data_cor1 <- cor(final_data)
corrplot(final_data_cor1, method="square", col = rev(colorRampPalette(c("#B40F20", "#FFFFFF", "#2E3A87"))(100)), type="lower", tl.col="black", tl.srt=60, tl.cex = 0.6)
```

## ggpairs plot 
The features selected are selected based on high correlation between factors in the correlation plot.
```{r}
ggpairs(final_data, columns = c("popularity", "danceability", "loudness", "instrumentalness"), lower = list(continuous = "smooth"), upper = list(continuous = "cor"))
```

## Standardised metrics
Popularity vs: danceability, energy, speechiness, acousticness, instrumentalness, liveness, valence
```{r}
basic_plots <- function(x){
  # plot without transparency  
  plot_nt <- ggplot(final_data, aes(x = !!sym(x), y = popularity)) +
    geom_point(alpha = 0.1)
  # plot with transparency 
  plot_wt <- ggplot(final_data, aes(x = !!sym(x), y = popularity)) +
    geom_bin2d(alpha = 0.7) +
    scale_fill_gradientn(colors = c("#440154", "#30678D", "#35B778", "#FDE724", "#FFFFFF"))
  # Return both plots 
  return(list(plot_nt, plot_wt))
}

metrics <- c('danceability', 'energy', 'speechiness', 'acousticness', 'instrumentalness', 'liveness', 'valence')
for (i in metrics) {
  plots <- basic_plots(i)
  grid.arrange(plots[[1]], plots[[2]], ncol = 2)
}
```

## Genre
```{r}
# Assign the genre name based on the dummy variables
get_genre_name <- function(x) {
  ifelse(x["two_genre"] == 1, "2_genres",
    ifelse(x["three_genre"] == 1, "3_genres",
      ifelse(x["four_genre"] == 1, "4_genres",
        ifelse(x["five_genre"] == 1, "5_genres",
          ifelse(x["rock"] == 1, "rock",
            ifelse(x["country"] == 1, "country",
              ifelse(x["jazz"] == 1, "jazz",
                ifelse(x["electronic"] == 1, "electronic",
                  ifelse(x["classical"] == 1, "classical",
                    ifelse(x["world"] == 1, "world",
                      ifelse(x["kids"] == 1, "kids",
                        ifelse(x["other"] == 1, "other",
                          ifelse(x["rap"] == 1, "rap", "pop")))))))))))))
}
# Apply the function to each row of the data frame and create a new column with the genre names
temp_data <- data.frame(final_data)
temp_data$genre_name <- apply(final_data[, -1], 1, get_genre_name)

# Create a bar plot of mean popularity by genre
mean_popularity <- tapply(temp_data$popularity, temp_data$genre_name, mean)
barplot(mean_popularity, xlab = "Genre", ylab = "Mean Popularity", col = "steelblue", main = "Mean Popularity by Genre", las = 2, cex.names = 0.8)
```

# Split data

```{r}
# Split the dataset into training and testing sets
set.seed(1)
split <- initial_split(final_data, prop = 0.7)
train <- training(split)
test <- testing(split)

X_train <- train[,-1]
y_train <- train$popularity

X_test <- test[,-1]
y_test <- test$popularity

```


# Baseline model

```{r}
#sum(complete.cases(final_data))
baseline <- lm(popularity ~ danceability, data = train)
predictions_baseline <- augment(baseline)

model_loess <- loess(popularity ~ danceability, span = 0.75, data = train)
predictions_loess <- augment(model_loess)

gm_scatterplot <- 
  ggplot(final_data, aes(x = danceability, y = popularity)) +
  geom_point()
gm_scatterplot

gm_scatterplot +
  geom_line(data = predictions_baseline, size = 1,
            color = "blue",
            linetype = "dashed",
            aes(x = danceability, y = .fitted)) +
  geom_line(data = predictions_loess, size = 1,
            color = "green",
            aes(x = danceability, y = .fitted)) 


#summary(baseline)
```

# Ridge and Lasso
```{r}
# gaussian is for quantitative outcome
ridge_fit <- glmnet(X_train, y_train,
                    family = "gaussian", 
                    alpha = 0)
lasso_fit <- glmnet(X_train, y_train,
                    family = "gaussian",
                    alpha = 1)
```

```{r}
lam <- 0.8
cbind(coef(ridge_fit, s = lam), coef(lasso_fit, s = lam)) # combining two columns together, the first column is ridge, and the second col is lasso
plot(ridge_fit, xvar = "lambda")
plot(lasso_fit, xvar = "lambda")
```

Training accuracy for ridge - mean square error (MSE)

```{r}
ridge_fit_predict <- predict(ridge_fit, 
        # type = "class", # "class is for classification
        newx = X_train,
        s = lam)
MSE_ridge <- mean((ridge_fit_predict - Y_train)^2)
MSE_ridge
```

Test accuracy for ridge - mean square error (MSE)

```{r}
ridge_fit_predict <-predict(ridge_fit, 
        # type = "class",
        newx = X_test,
        s = lam)   
MSE_ridge <- mean((ridge_fit_predict - Y_test)^2)
MSE_ridge
```


